---
title: "FRS Dementia Study"
author: "Pallavi Malladi"
date: "Winter 2019"
output:
  html_document:
    keep_md: yes
    theme: spacelab
---
## Goal: Determine which ethnicity and geographic code have a higher inclincation for developing dementia. Is there any correlation between the ethnicity and geography?

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
dementia_data <- readr::read_csv("~/Desktop/FRS_417/pallavimalladi/Dementia_Study/Recorded_Dementia_Diagnoses_by_Ethnicity.csv")
```

```{r}
names(dementia_data)
```

Rename some of the variables. Notice that I am replacing the old `dementia_data` data.
```{r}
dementia_data <- 
  dementia_data %>% 
  dplyr::rename(
          Recorded_Date = 'ach_date',
          Organization_Code  = 'COMMISSIONER_ORGANISATION_CODE',
          Clinical_Commisioning_Groups  = 'CCG_NAME',
          Ethnicity  = 'Measure',
          Value   = 'Value'
          )
dementia_data
```

## Explore the data using the method that you prefer. Below, I am using a new package called `skimr`. If you want to use this, make sure that it is installed.
```{r}
install.packages("skimr")
##installs a new package (not the tidyverse!)
```

```{r}
library("skimr")
```

```{r}
dementia_data %>% 
  skimr::skim()
##skimr is like another summary function
```

## 2. Run the code below. Are there any NA's in the data? Does this seem likely?
```{r}
dementia_data %>% 
  summarize(number_nas= sum(is.na(dementia_data)))
## Technically, there are no NA's, but it is possible that there is another value such as -999 that represents a null value
```

##Produce a summary that shows the number of observations by ethnicity.
```{r}
names(dementia_data) ##I'm using this to get a general idea of the column headers so I know which notation to use within my code
```

```{r}
dementia_data %>% 
  group_by(Ethnicity) %>% ##groups by taxonomic order
  summarize(n())## n represents the number of observations; each order has a certain number of observations
```

## 5. Mammals have a range of life histories, including lifespan. Produce a summary of lifespan in years by order. Be sure to include the minimum, maximum, mean, standard deviation, and total n.
```{r}
dementia_data <-dementia_data %>%
  ## MAKE CODE TO REMOVE THE NOT_DEFINED ETHNICITY SINCE IT IS NOT TELLING US RELEVANT INFORMATION
  arrange(desc(Value))
dementia_data
```

## 6. Let's look closely at gestation and newborns. Summarize the mean gestation, newborn mass, and weaning mass by order. Add a new column that shows mean gestation in years and sort in descending order. Which group has the longest mean gestation? What is the common name for these mammals?
```{r}
life_history %>% 
  group_by(order) %>% 
  summarize(mean_gestation=mean(gestation, na.rm=TRUE), ##summarizes mean for specified values
            mean_newborn_mass=mean(newborn, na.rm=TRUE),
            mean_wean_mass=mean(wean_mass, na.rm=TRUE)) %>% 
  arrange(desc(mean_gestation)) 
```
## Proboscidea has longest mean gestation; "The Proboscidea (from the Greek προβοσκίς and the Latin proboscis) are a taxonomic order of afrotherian mammals containing one living family (Elephantidae) and several extinct families"
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

