---
title: "FRS Dementia Study"
author: "Pallavi Malladi and Tiffany Sin"
date: "Winter 2019"
output: ioslides_presentation
---
## Introduction: For our presentation, we wanted to focus our attention on data in the health field. We chose to analyze data regarding dementia in England and we got our data from NHS Digital, which publishes the recorded dementia diagnoses each month with the latest from January 2019. This data was collected because not everyone gets a formal diagnosis for dementia. The data compares people who were thought to have dementia and those who were acutally diagnosed. These people were all 65 and older.

## Goal: Determine which Race and geographic code have a higher inclincation for developing dementia. Is there any correlation between the Race and geography?

## The data is collected between April 2018 and January 2019. Data was collected at the end of each month, and recorded values represent the number of people who showed signs/symptoms of Dementia at various clinics throughout England.

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r}
dementia_data <- readr::read_csv("~/Desktop/FRS_417/pallavimalladi/Dementia_Study/Recorded_Dementia_Diagnoses_by_Race.csv")
```

```{r}
names(dementia_data)
```

## 2. Run the code below. Are there any NA's in the data? Does this seem likely?
```{r}
dementia_data %>% 
  summarize(number_nas= sum(is.na(dementia_data)))
## Technically, there are no NA's, but it is possible that there is another value such as -999 that represents a null value
```

## Produce a summary that shows the number of observations by Race.
```{r}
names(dementia_data) ##I'm using this to get a general idea of the column headers so I know which notation to use within my code
```

Rename some of the variables. Notice that I am replacing the old `dementia_data` data.
```{r}
dementia_data <- 
  dementia_data %>% 
  dplyr::rename(
          Recorded_Date = 'ach_date',
          Organization_Code  = 'COMMISSIONER_ORGANISATION_CODE',
          Clinical_Commisioning_Groups  = 'CCG_NAME',
          Race  = 'Measure',
          Dementia_Rate_Indicator   = 'Value'
          )
dementia_data
```

```{r}
dementia_data %>% 
  group_by(Race) %>% ##groups by taxonomic order
  summarize(n())## n represents the number of observations; each order has a certain number of observations

## From this line of code, we see that each Race has 1950 observations
```

## Rename the data values for Race so that the names are not so long (long names may cause problems in later lines of code, especially when trying to make plots)
```{r}
dementia_data<- dementia_data %>% 
  separate(Race, into= c("Race","Race2","Race3","Race4","Race5","Race6","Race7", "Race8" ), sep = "_") %>% 
  select(-Race2, -Race3,-Race4,-Race5,-Race6, -Race7,-Race8)
dementia_data
##Please not that "NOT" means not defined
```

## Mutate the Dementia_data set to not include the "NOT" (NOT DEFINED) since it is not contributing to our results dealing with specific Races.
```{r}
dementia_data<- dementia_data %>% 
  filter(Race!= "NOT")
dementia_data
```

## Rearrange the data table to show decreasing order of Dementia Indicator. This will be a good indicator as to the highest race coming into clinics with Dementia diagnoses
```{r}
dementia_data <-dementia_data %>%
  arrange(desc(Dementia_Rate_Indicator))
dementia_data
```

##Find the total number of people coming into the NIH clinics for each Race between April 2018 to January 2019
```{r}
dementia_data %>% 
  group_by(Race) %>% 
  summarize(total= sum(Dementia_Rate_Indicator)) %>% 
  arrange(desc(total))
## total represents number of people who have come into clinics within each race
##Notice how from this data, the White and Asian Races have the highest Dementia Diagnoses, while Black data is closest to 0. This should be very apparent in our upcoming plots.
```

## Making a plot to visually show the difference between. Plot is shown for each month and shows barplot comparing Races showing Dementia
```{r}
dementia_data %>% 
  ggplot(aes(x=Race, y=Dementia_Rate_Indicator, fill=Race))+
  geom_bar(stat="identity")+
  theme(plot.title = element_text(size = rel(1.95)))+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust=1))
```

## Find the Clincal Commisioning Group where the Dementia_Rate_Indicator is highest for each Race. Use this information to find geographical location where dementia is prevalent for that specific Race.
```{r}
dementia_data %>% 
  group_by(Clinical_Commisioning_Groups) %>% 
  summarize(total= sum(Dementia_Rate_Indicator)) %>% 
  arrange(desc(total)) 
## total represents number of people who have come into each specific clinic
```

## Making the plot showing the basic race data for the top 5 clinical commisioning groups with the most amount of clinic visits
```{r}
CCG_data <- dementia_data %>% 
  select (Clinical_Commisioning_Groups, Race, Dementia_Rate_Indicator) %>% 
  filter(Clinical_Commisioning_Groups=="NHS Leeds CCG"| Clinical_Commisioning_Groups=="NHS Birmingham and Solihull CCG"| Clinical_Commisioning_Groups=="NHS Mid Essex CCG"| Clinical_Commisioning_Groups=="NHS Bedfordshire CCG" | Clinical_Commisioning_Groups=="NHS Leicester City CCG")
CCG_data
```

```{r}
CCG_data %>% 
  ggplot(aes(x=Race, y= Dementia_Rate_Indicator, fill=Race))+
  geom_bar(stat="identity") +
  facet_wrap(~Clinical_Commisioning_Groups)+
  theme(plot.title = element_text(size = rel(1.95)))+
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 60, hjust=1))
```

```{r}
general_data<- dementia_data %>% 
  select(Organization_Code, GEOGRAPHY_CODE,Clinical_Commisioning_Groups, Race, Dementia_Rate_Indicator)
general_data
```

## Conclusion: There seems to be a significantly higher amount of Whites coming into the clinics. This may indicate a social difference between White and other Races. For example, other Races may have cultural backgrounds that stigmatizes mental health or going to hospitals regularly for chekcups. We were not sue how to work around this significantly higher data, but it does bring about an important topic for discussion regarding mental health, social taboos, financial burders, and personal care.

## Facts from the study: 464,413 people had a diagnosis of demenia as of January 31st 2019. Camden CCG has the highest estimated dementia diagnosis rate, at 90.60%. Kernow CCG has the lowest estimated dementia diagnosis rate, at 52.10%